Temporal stability in human interaction networks

### 2015-11-03 paper.tex
incorporada revisao do chu

	\item Developer's list for the standard C++ library\footnote{gmane.comp.gcc.libstdc++.devel is list ID in Gmane.}, with computer programmers from different countries. English is the prevailing language. Abbreviated as CPP from now on.
	\item List of the MetaReciclagem project\footnote{gmane.politics.organizations.metareciclagem is list ID in Gmane.}, a Brazilian email list for digital culture. 	Portuguese is the prevailing language, although some messages are writte in Spanish and English. Abbreviated as MET from now on.
\end{itemize}

@@

																																																																																%\subsubsection{Sectioning of networks in peripheral, intermediary and hubs sectors}\label{sectioning}
																																																																																\subsection{Principal Component Analysis of topological metrics}\label{sec:pca}
																																																																																Principal Component Analysis (PCA) is a well documented technique~\cite{pca}
																																																																																and was employed in the investigation to acquire knowledge about:
																																																																																1) which metrics contribute to each principal component and in what proportion;
																																																																																2) how much of the dispersion is concentrated in each component;
																																																																																3) expected values and dispersions for these quantities over various networks.
																																																																																This enables a characterization of the human interaction networks
																																																																																in terms of the relative importance of networks measures and the way they combine.

																																																																																Let $\mathbf{X}=\{X[i,j]\}$ be a matrix where each element is the value
																																																																																of each metric $j$ at vertex $i$ .
																																																																																Let
																																																																																$\mu_X [j]=\frac{\sum_i X[i,j]}{I}$ be the mean of metric $j$ over all $I$ vertices, 
																																																																																$\sigma_X [j]=\sqrt{\frac{\sum_i (X[i,j]-\mu_X [j])^2}{I}}$ the standard deviation of metric $j$,
																																																																																and $\mathbf{X'}=\{X'[i,j]\}=\left\{\frac{X[i,j]-\mu_X[j]}{\sigma_X[j]}\right\}$ 
																																																																																the matrix with the \emph{z-score} of each metric. 
																																																																																Let $\mathbf{V}=\{V[j,k]\}$ be the matrix $J\times J$ of eigenvectors
																																																																																of the covariance matrix $\mathbf{C}$
																																																																																of $\mathbf{X'}$, one eigenvector per column.
																																																																																Each eigenvector combines the original measures into one principal component, therefore
																																																																																$V'[j,k]=100\frac{|V[j,k]|}{\sum_{j'} |V[j',k]|}$
																																																																																is the percentage of the principal component $k$
																																																																																that is proportional to the metric $j$.
																																																																																%With $k$ eigenvectors 
																																																																																%$D[k]$,
																																																																																%it is enough to 
																																																																																Let $\mathbf{D}=\{D[k]\}$ be the eigenvalues associated with the eigenvectors $\mathbf{V}$,
																																																																																then $D'[k]=100\frac{D[k]}{\sum_{k'}D[k']}$
																																																																																is the percentage of total dispersion of the system that the principal component $k$
																																																																																is responsible for.
																																																																																We consider, in general, the three greatest eigenvalues and
																																																																																the respective eigenvectors in percentages:
																																																																																$\{(D'[k],\;V'[j,k])\}$.
																																																																																These usually sum up between 60 and 95\% of the dispersion
																																																																																and reveal patterns for a first analysis.
																																																																																In particular, 
																																																																																given $L$ snapshots $l$ of the interaction network,
																																																																																we are interested in the mean
																																																																																$\mu_{V'}[j,k]$
																																																																																and the standard deviation $\sigma_{V'}[j,k]$ 
																																																																																of the contribution of metric $j$ to the principal component $k$,
																																																																																and the mean
																																																																																$\mu_{D'}[k]$
																																																																																and the standard deviation 
																																																																																$\sigma_{D'}[k]$
																																																																																of the contribution of the component $k$ to the dispersion
																																																																																of the system:

@@

																																																																																%This is a way to enhance the reliability of the methods,
																																																																																%of algorithmic routines, of data consistency,
																																																																																%and of the results themselves.
																																																																																%Also, we believe that this practice raises the scientific contribution of the work,
																																																																																%handing not only the framework and the results, 
																																																																																%but the exact data and processes that render them~\cite{openSci}.
																																																																																\section{Results and discussion}\label{sec:results}
																																																																																%Remarkable features from the analysis of the four email lists are:
																																																																																%\begin{itemize}
																																																																																%    \item The activity along time is practically the same for all lists, thus suggesting stable patterns.
																																																																																%    \item The fraction of participants in each Erd\"os sector is stable along time and can be determined even with very few messages 
																																																																																%    \item The topological metrics combine into principal components in PCA in the same way for all lists and all snapshots (). 
																																																																																%        \item Symmetry measures of the topology, as defined in this article, present more dispersion than the usual clustering coefficient (Section~\ref{prevalence}).
																																																																																%    \item Typology speculations are immediate from results (Section~\ref{sec:pty}).
																																																																																%\end{itemize}
																																																																																\subsection{Activity along time}\label{constDisc}
																																																																																Regular patterns of activity were observed along time
																																																																																in the scales of seconds, minutes, hours, days and months.
																																																																																Histograms in each of the time scales were computed as were circular average and dispersion values, and The results are given in Tables~\ref{tab:circ}-\ref{tab:min2}. For example, uniform activity is found with respect to seconds, minutes and days of the months. Weekend days exhibit about half the activity of regular weekdays, and there is a peak of activity between 11am and 12pm.

@@

																																																																																\begin{table}
																																																																																	\caption{Activity percentages on months along the year. 	Activity is usually concentrated in Jun-Aug and/or in Dec-Mar, potentially due to academic calendars, vacations and end-of-year holidays. This table corresponds to activity in LAU. Similar results are shown in Section~\ref{si:months} of the Supporting Information document.}
																																																																																	\footnotesize
																																																																																	\input{tables/tabMonthsLAU}
																																																																																	\label{tab:min2}
																																																																																\end{table}

>>>>

\item Developer's list for the standard C++ library\footnote{gmane.comp.gcc.libstdc++.devel is list ID in Gmane.}, with computer programmers from different countries. English is the prevailing language. Abbreviated as CPP from now on.
\item List of the MetaReciclagem project\footnote{gmane.politics.organizations.metareciclagem is list ID in Gmane.}, a Brazilian email list for digital culture. 	Portuguese is the prevailing language, although some messages are written in Spanish and English. Abbreviated as MET from now on.
\end{itemize}

@@

%\subsubsection{Sectioning of networks in peripheral, intermediary and hubs sectors}\label{sectioning}
\subsection{Principal Component Analysis of topological metrics}\label{sec:pca}
Principal Component Analysis (PCA) is a well documented technique~\cite{pca}, used here to address the following questions:	1) which metrics contribute to each principal component and in what proportion;	2) how much of the dispersion is concentrated in each component;	3) which are the expected values and dispersions for these quantities over various networks.	This enables one to characterize human interaction networks in terms of the relative importance of network metrics and the way they combine.

Let $\mathbf{X}=\{X[i,j]\}$ be a matrix where each element is the value
of each metric $j$ at vertex $i$ .
Let
$\mu_X [j]=\frac{\sum_i X[i,j]}{I}$ be the mean of metric $j$ over all $I$ vertices, 
$\sigma_X [j]=\sqrt{\frac{\sum_i (X[i,j]-\mu_X [j])^2}{I}}$ the standard deviation of metric $j$,
and $\mathbf{X'}=\{X'[i,j]\}=\left\{\frac{X[i,j]-\mu_X[j]}{\sigma_X[j]}\right\}$ 
the matrix with the \emph{z-score} of each metric. 
Let $\mathbf{V}=\{V[j,k]\}$ be the matrix $J\times J$ of eigenvectors
of the covariance matrix $\mathbf{C}$
of $\mathbf{X'}$, one eigenvector per column.
Each eigenvector combines the original measures into one principal component, therefore
$V'[j,k]=100\frac{|V[j,k]|}{\sum_{j'} |V[j',k]|}$
is the percentage of the principal component $k$
that is proportional to the metric $j$.
%With $k$ eigenvectors 
%$D[k]$,
%it is enough to 
Let $\mathbf{D}=\{D[k]\}$ be the eigenvalues associated with the eigenvectors $\mathbf{V}$,
then $D'[k]=100\frac{D[k]}{\sum_{k'}D[k']}$
is the percentage of total dispersion of the system that the principal component $k$
is responsible for.
We consider, in general, the three greatest eigenvalues and
the respective eigenvectors in percentages:
$\{(D'[k],\;V'[j,k])\}$.
These usually sum up between 60 and 95\% of the dispersion
and reveal patterns for a first analysis.
In particular, 
given $L$ snapshots $l$ of the interaction network,
we are interested in the mean
$\mu_{V'}[j,k]$
and the standard deviation $\sigma_{V'}[j,k]$ 
of the contribution of metric $j$ to the principal component $k$,
and the mean
$\mu_{D'}[k]$
and the standard deviation 
$\sigma_{D'}[k]$
of the contribution of the component $k$ to the dispersion
of the system:

@@

%This is a way to enhance the reliability of the methods,
%of algorithmic routines, of data consistency,
%and of the results themselves.
%Also, we believe that this practice raises the scientific contribution of the work,
%handing not only the framework and the results, 
%but the exact data and processes that render them~\cite{openSci}.
\section{Results and discussion}\label{sec:results}
%Remarkable features from the analysis of the four email lists are:
%\begin{itemize}
%    \item The activity along time is practically the same for all lists, thus suggesting stable patterns.
%    \item The fraction of participants in each Erd\"os sector is stable along time and can be determined even with very few messages 
%    \item The topological metrics combine into principal components in PCA in the same way for all lists and all snapshots (). 
%        \item Symmetry measures of the topology, as defined in this article, present more dispersion than the usual clustering coefficient (Section~\ref{prevalence}).
%    \item Typology speculations are immediate from results (Section~\ref{sec:pty}).
%\end{itemize}
\subsection{Activity along time}\label{constDisc}
Regular patterns of activity were observed along time
in the scales of seconds, minutes, hours, days and months.
Histograms in each of the time scales were computed as were circular average and dispersion values, and the results are given in Tables~\ref{tab:circ}-\ref{tab:min2}. For example, uniform activity is found with respect to seconds, minutes and days of the months. Weekend days exhibit about half the activity of regular weekdays, and there is a peak of activity between 11am and 12pm.

@@

\begin{table}
\caption{Activity percentages on months along the year. 	Activity is usually concentrated in Jun-Aug and/or in Dec-Mar, potentially due to academic calendars, vacations and end-of-year holidays. This table corresponds to activity in LAU. Similar results are shown for other lists in Section~\ref{si:months} of the Supporting Information document.}
\footnotesize
\input{tables/tabMonthsLAU}
\label{tab:min2}
\end{table}
